---
title: "Social_Survey_And_Lakers"
author: "Mia Fitzgerald"
date: "4/7/2026"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(tidyverse)
library(lubridate)
library(knitr)
library(forcats)

```


### Combining Levels


```{r}
gss1 <- gss_cat %>%
  mutate(marital = fct_recode(marital, 
"Pre-Marriage" = "Never married",
"Marriage" = "Married",
"Marriage" = "Separated",
"Post-Marriage" = "Divorced",
"Post-Marriage" = "Widowed",
"Missing" = "No answer"))

gss1 %>%
  count(marital) %>%
  kable()

```


I consolidated the specific marital statuses in the gss_cat dataset into four categories using fct_recode(). By grouping "Never married" as Pre-Marriage, "Married" and "Separated" as Marriage, and "Divorced" and "Widowed" as Post-Marriage. Finally, it generates a count of these new groups and displays them in a table.
^[Question 1 complete]


### Proportion of Responses in Each Year


```{r}
gss2 <- gss1 %>%
  count(year, marital) %>%
  group_by(year) %>%
  mutate(prop = n/sum(n))

gss2 %>%
  kable()

```


First, it counts how many people fall into each marital category for every year. Then, it groups the data by year so that it can calculate the portion of that specific year's total. The final result is displayed as a formatted table.
^[Question 2 complete]


### Status Reordered by Year


```{r}
gss3 <- gss2 %>%
  group_by(year, marital) %>%
  arrange(desc(year), desc(prop)) %>%
  ungroup()

gss3 %>%
  kable()

```


I group the data so I can logically rank the information. The arrange() function puts the most recent years at the top and sorts status from the highest to the lowest. Making sure that for every year, you immediately see the most common marital status first. Finally, it removes the grouping and displays the results in a table.
^[Question 3 complete]


### Proportion Versus Year


```{r}
ggplot(gss3, aes(x = year, y = prop, color = marital)) + 
  geom_point() +
  geom_line()

```


^[Question 4 complete]


### Integer to Date Error


```{r}
lakers %>%
  mutate(new_date = make_date(date)) %>%
  select(date, new_date, time, player, result, etype, points) %>%
  head(n=6) %>%
  kable()

```


^[Question 5 complete]


### Subset of Dataset Including Parsed Dates


```{r}
lakers2 <- lakers %>%
  mutate(date2 = as.character(date)) %>%
  mutate(date2 = ymd(date2)) 

lakers2 %>%
  select(date, date2, team, etype, time, points) %>%
  head(6) %>%
  kable()

```


^[Question 6 complete]


### Games Played on Different Days of the Week


```{r}
lakers_bar <- lakers2 %>%
  distinct(date2) %>%
  mutate(day = wday(date2, label = TRUE, abbr = TRUE))

ggplot(lakers_bar, aes(day)) +
  geom_bar()

```


^[Question 7 complete]